### 2025-02-21 code/processing-code/processingfile.qmd
Data cleaning

---
title: "Data cleaning demo"
author: "Guozheng Yang"
date: "2024-02-07"
output: html_document
---

# Processing script

This contains the same code and comments/information as `processingcode.R`.

But one can directly see some of the outputs.

# Setup

Load needed packages.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

# Data loading

```{r}
# path to data
data_location <- here::here("data","raw-data","data0.xlsx")
rawdata <- readxl::read_excel(data_location)
```

# Check data

The codebook for this data is in the READMD file in the raw-data sub-folder.

An overview of the data structure.

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

# Cleaning

I will check the variables one by one.

*ID*: looks good

*Sex*: looks good

*Age*: looks good

*Height*: looks good

*Weight*: looks good

*BMI*: looks good

*TBW_L & TBW_pct*:

*TBW_L* only makes sense when it is lower than body weight. 
The *TBW_pct* should be within (0, 1).
Wrongly measured *TBW_L* will also result in weird read in other body indices.
Note that all individuals with *TBW_pct*>73% have negative *Fat_kg*.
Also, all individuals with *TBW_pct*<0 have *Fat_pct*>1. 
Here I keep all *TBW_pct*>73% or *TBW_pct*<0 and corresponding *TBW_L* as missing.
Other body composition indices will also be kept as NA.

```{r}
d1 <- rawdata %>%
  mutate(across(TBW_L:FFMI, ~ ifelse(TBW_pct > 0.73 | TBW_pct < 0, NA, .))) %>%
  mutate(across(ends_with("_pct"), ~ .*100)) # multiply percentages by 100
```

*Leptin_ng_ml*: looks good (one extremely high - 44+, but it's still reasonable)

*CRP_mg_dl*: looks good

*INF_Gamma_pg_ml*: looks good

*TNF_Alpha_pg_ml*: looks good

*IL_10_pg_mL*: looks good

*CD4_plus*: looks good

*CD4_plus_pct*: looks good (multiplied by 100 in last step)

*CD8_plus*: looks good

*CD8_plus_pct*: looks good (multiplied by 100 in last step)

All the other variables are fine. But I will round CD4+ counts and CD8+ counts to the nearest integer.

```{r}
d2 <- d1 %>%
  mutate(CD4_plus=round(CD4_plus),
         CD8_plus=round(CD8_plus))
```

The data cleaning process is completed. Let's take a final look at the cleaned data.

```{r}
dplyr::glimpse(d2)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```