### 2025-04-28 Review - Grismer/yang-review-byhopeg.qmd
Review - Grismer

---
title: Project Review Template 
author: Hope Grismer
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Sex-Based Differences in Tuberculosis Immune Response: The Role of Leptin

Name of project author(s): G. Yang

Name of project reviewer: Hope Grismer

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background is strong and clear, placing the project into the broader context of tuberculosis, sex differences in immune responses, and the metabolic role of leptin.\
The motivation for the study is well stated, though it would be even stronger if the introduction included more direct references or citations supporting the specific immune differences between sexes (besides leptin pathways).

Suggestion: If no direct references exist, explicitly framing this as a hypothesis would improve clarity.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

-   strong contextualization and motivation

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The research questions and hypotheses are clearly articulated:

-   Women may have stronger immune responses during TB infection due to higher body fat/leptin.

-   Leptin mediates the link between body composition and immune activation.

They are closely aligned with the data being collected and analyzed.

### Summary assessment

-   question/hypotheses fully clear

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data sources and measurements are very well explained. The manuscript is transparent about how all variables were generated, and methods are described sufficiently to replicate measurements.

### Summary assessment

-   source and overall structure of data well explained

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The exploratory analysis in the manuscript is strong, with thoughtful visualizations and descriptive statistics.

However, data wrangling (cleaning) is barely described in the manuscript itself.\
From the processing script, it is clear that minimal cleaning was necessary — but that context should still be briefly mentioned in the manuscript or supplemental materials.\
Also, it would be better practice to have the exploratory plots in a separate EDA script (instead of combining all statistical analysis into one giant `.R` file).

### Summary assessment

-   some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Minor suggestion: Including a multivariable linear model adjusting for possible confounders (e.g., age, sex, body composition) could have strengthened the findings, especially for leptin prediction.

But overall, analysis choices are sound and competently implemented.

### Summary assessment

-   strong and reasonable analysis

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The figures and tables are clear and easy to interpret.

However, a few suggestions:

-   Table 2 is confusing because it combines linear and Poisson regression coefficients.\
    ➔ Better to split into two separate tables.

-   P-values should be included alongside 95% confidence intervals for clarity.

-   Explain significance thresholds (e.g., p \< 0.05) explicitly in captions or footnotes.Summary assessment

<!-- -->

-   results are very well presented (tiny room for improvement)

### Summary assessment

-   results are presented ok, with room for improvement

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion is strong, thoughtful, and places findings in a wider biological and public health context.

The link between findings and broader TB immunology research is made very well.

Minor suggestion: Future research directions could be elaborated slightly more (e.g., larger cohorts, inclusion of HIV+ individuals).

### Summary assessment

-   strong, complete and clear discussion

## Further comments

Slightly better description of cleaning/EDA would make it even stronger.

# Overall project content evaluation

Evaluate overall features of the project by filling in the sections below.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

-   Three main Quarto files corresponding to cleaning, EDA, and analysis

-   No junk files

-   Clear naming conventions

### Summary assessment

-   Well structured

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The manuscript itself is very well documented.\
The code is mostly documented, but:

-   Some sections lack comments explaining the purpose of each chunk

-   More small comments explaining model fitting choices would improve readability.

### Summary assessment

-   decently documented with some gaps

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

-   All code files are provided and it is reproducible

### Summary assessment

-   fully reproducible without issues

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The project is thorough with:

-   Mediation models

-   Machine learning prediction models

-   A simulation study to validate methods

However, alternative regression models (e.g., multivariable modeling adjusting for covariates) could have been explored a bit more.

### Summary assessment

-   strong level of thorougness (with minor suggestions)

## Further comments

Thank you for an amazing project! I really was inspired in getting to look yours over. It was super well-rounded project– very close to publication-level. I especially liked the depth and strength of writing in your manuscript.